---
title: "R practice ITS"
author: "Dan Weinberger"
date: '2023-04-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(MASS)
library(ggplot2)
library(readxl)
```

The 10-valent pneumococcal vaccine was introduced in country X in August 2010

```{r}
# a1 <- read_excel('Data/raw_data_u5.xlsx') %>%
#   filter(!is.na(DOB) & DOB>='2003-01-01') %>%
#   rename(death_date=`Date of death`, icd1=`Cause of death`) %>%
#   mutate(aged=as.numeric(difftime(death_date, DOB, units = "days")),
#          agem=round(aged/30.3)) %>%
#   select(agem, death_date, icd1 ) %>%
# filter(agem>=2)
# saveRDS(a1,'./Data/line_list_formatted.rds')
```

```{r}
a1 <- readRDS('./Data/line_list_formatted.rds') %>%
  filter(death_date>='2005-01-01')
```

## Part 1: Prepare your data

### Step 0: Create a histogram of the age distribution, check the range of ages and dates

```{r}

```


### Step 1: Flag deaths in the range J12-J18, as well as non-respiratory deaths

```{r}

```


### Step 2: Create a time series for J12-J18 and all-non-respiratory deaths

```{r}

```


### Step 3: Plot your time series for J12:J18 and J12_J18/all-non-respiratory

```{r}

```

## Part 2: ITS

### Step 1: Create variables for pre/post periods 
-what are the time periods you will use?
- Create an index variable to capture trends
- Create a monthly factor variable OR sine and cosine terms for seasonality
-Create binary variables for pre/post periods
-Create linear splines to allow trend to change at hinge points

ESCAPE HATCH 1: in case you haven't succeeded in creating a time series, start from here
```{r, eval=F}
ts1 <- readRDS( './Data/formatted_ts.rds')
```

```{r}

```

### Step 2: Fit and compare models

ESCAPE HATCH 2: In case you don't have the variables needed for the ITS analysis, start here
```{r}
 ts1 <- readRDS('./Data/formatted_ts_with_its.rds')

```

Will you use Poisson or negative binomial models here?

Model 1: Trend +seasonality + post vaccine binary variables +offset(ACM_noJ)
```{r}

```

Model 2: Trend +seasonality + post vaccine splines +offset(ACM_noJ)

```{r}

```

Model 3: seasonality + post vaccine binary variables+offset(ACM_noJ)

```{r}

```

Model 4: seasonality + post vaccine splines+offset(ACM_noJ)

```{r}

```

### Step 3: Calculate Fitted, counterfactual, and Rate ratio for best model

Generate fitted value using predict()
```{r}

```

Generate counterfactual fitted values by zeroing out the vaccine effects

```{r}

```

add these to your plot of observed values

```{r}
```

Calculate the rate ratio for the last time point
```{r}

```

Calculate lives saved during the post-vaccine period


