---
title: "R practice ITS"
author: "Dan Weinberger"
date: '2023-04-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(ggplot2)
library(readxl)
```

The 10-valent pneumococcal conjugate vaccine was introduced in country X in August 2010

```{r}
# a1 <- read_excel('Data/raw_data_u5.xlsx') %>%
#   filter(!is.na(DOB) & DOB>='2003-01-01') %>%
#   rename(death_date=`Date of death`, icd1=`Cause of death`) %>%
#   mutate(aged=as.numeric(difftime(death_date, DOB, units = "days")),
#          agem=round(aged/30.3)) %>%
#   select(agem, death_date, icd1 ) %>%
# filter(agem>=2)
# saveRDS(a1,'./Data/line_list_formatted.rds')
```

```{r}
a1 <- readRDS('../Data/line_list_formatted.rds')
```

## Part 1: Prepare your data

### Step 0: Create a histogram of the age distribution, check the range of ages and dates

```{r}
hist(a1$agem)

range(a1$agem)
```


### Step 1: Flag deaths in the range J12-J18

```{r}

a1 <- a1 %>%
  mutate( J12_J18 = if_else( icd1 %in% c('J12','J13','J14','J15','J16','J17','J18'),1,0),
          acm_noJ = 1-(substr(icd1,1,1)=='J') 
          )


```


### Step 2: Create a time series for J12-J18 and all-cause deaths

```{r}

ts1 <- a1 %>%
  mutate(date=as.Date(floor_date(death_date, 'month'))) %>%
  group_by(date) %>%
  summarize(J12_J18=sum(J12_J18), acm_noJ= sum(acm_noJ)) %>%
  filter(date>='2005-01-01')

saveRDS(ts1, '../Data/formatted_ts.rds')
```


### Step 3: Plot your time series for J12:J18 and J12_J18/all-cause

```{r}

p1 <- ggplot(ts1, aes(x=date, y=J12_J18) ) +
  geom_line() +
  theme_classic() +
  ylim(0, NA)
p1
```

## Part 2: ITS

### Step 1: Create variables for pre/post periods 
-what are the time periods you will use?
- Create an index variable to capture trends
- Create a monthly factor variable OR sine and cosine terms for seasonality
-Create binary variables for pre/post periods
-Create linear splines to allow trend to change at hinge points

```{r}

 post1_index= which(ts1$date=='2010-08-01')
 post2_index= which(ts1$date=='2011-08-01')

ts1 <- ts1 %>%
  arrange(date) %>%
  mutate(index=row_number(),
         month=as.factor(month(date)),
         sin12=sin(2*pi*index/12),
         cos12=cos(2*pi*index/12),
         
         post1 = if_else(date>='2010-08-01',1,0),
         post2 = if_else(date>='2011-08-01',1,0),
         
                 spl1 = if_else(date>='2010-08-01', index-post1_index+1,0),
         spl2 = if_else(date>='2011-08-01', index-post2_index+1,0)
                  
    )
  
  saveRDS(ts1, '../Data/formatted_ts_with_its.rds')

```


### Step 2: Fit and compare models

Will you use Poisson or negative binomial models here?

Model 1: Trend +seasonality + post vaccine binary variables+offset(ACM)
```{r}
mod1 <- glm(J12_J18 ~ month +index +post1 +post2  + offset(log(acm_noJ)), data=ts1, family='poisson')

summary(mod1)
```

Model 2: Trend +seasonality + post vaccine splines+offset(ACM)

```{r}
mod2 <- glm(J12_J18 ~ month +index +spl1 +spl2 + offset(log(acm_noJ)), data=ts1, family='poisson')

summary(mod2)

```

Model 3: seasonality + post vaccine binary variables+offset(ACM)

```{r}
mod3 <- glm(J12_J18 ~ month + post1 +post2 + offset(log(acm_noJ)), data=ts1, family='poisson')

summary(mod3)

```

Model 4: seasonality + post vaccine splines +offset(ACM)

```{r}
mod4 <- glm(J12_J18 ~ month + spl1 +spl2 + offset(log(acm_noJ)), data=ts1, family='poisson')

summary(mod4)

```
```{r}
mod5 <- glm(J12_J18 ~ month +index +post1 +post2  + index*post1 + index*post2 + offset(log(acm_noJ)), data=ts1, family='poisson')

summary(mod5)
```
Compare the AIC scores

```{r}
AIC(mod1,mod2, mod3, mod4, mod5)
```




### Step 3: Calculate Fitted, counterfactual, and Rate ratio for best model

Generate fitted value using predict()
```{r}
ts1$pred1 <- predict(mod1, type='response')
```

Generate counterfactual fitted values by zeroing out the vaccine effects

```{r}
ts1.counter <- ts1 %>%
  mutate(post1=0,
         post2=0,
         spl1=0,
         spl2=0)

ts1$pred1.cf <- predict(mod1, newdata=ts1.counter, type='response')


```

```{r}
 p1 +
  geom_line(data=ts1, aes(x=date, y=pred1, color='red')) +
    geom_line(data=ts1, aes(x=date, y=pred1.cf, color='blue'))
```


Calculate the rate ratio for the last time point
```{r}
ts1 <- ts1 %>%
  mutate(RR =pred1/pred1.cf,
              deaths_averted=pred1.cf-pred1)

ts1$RR[nrow(ts1)]

plot(ts1$RR)
```

Calculate lives saved during the post-vaccine period

```{r}
ts1 %>%
  summarize(deaths_averted= round(sum(deaths_averted),-2))
```

